Experience Report: Haskell in Computational Biology

### 2012-03-08 mrfy_experience_report.tex
add something to 'what you must know to succeed' and filled in a bit of 'evaluation'

\subsection{What you must know to succeed}
- Ecosystem is loaded with tools that look promising but either are no longer maintained or are not ready for prime time (Pads, Backtrace)

\subsection{Evaluation}
- What does the local expert do? Do you have to have one?

- To the FP community: Those who want us to abandon our old tools for debugging imperative code have not provided a clear path to follow (both intellectually, and via software tools).

- No obvious body of knowledge on code improvement.

- The ecosystem contains tools that look promising and work great: Data.Memocombinators, Quickcheck, Parallel Strategies.
How to tell what's good? How did we discover what's good?
- Profiler can produce backtraces, but this was hard to discover.

>>>>

\subsection{What you must know to succeed}
\remark{I don't like this. I really struggled with this section. ---AG}
If others would like to follow in our footsteps, there are a few
things we believe one must know in order to succeed.
Firstly, some experience with functional programming concepts is a necessity. 
Specifically, coming from an imperative background, writing algorithms and 
modeling data without the ability to mutate state were the most difficult 
concepts to adapt to.
NEED MORE.

\subsection{Evaluation}
\remark{I think we need more in this section. ---AG}
In hindsight, we feel that there are several key facets of our project in 
Haskell that deserve evaluation.
Firstly, we found that existing tools and libraries were plentiful and 
typically helped us greatly.
In particular, Data.Memocombinators made dynamic 
programming with Viterbi exquisitely simple and elegant.
Control.Parallel.Strategies also made the parallelization of Viterbi trivial.
Most of the tools we used came straight from Hackage, and although 
documentation of some form was almost always present, it was sometimes 
difficult to tell if a particular tool was right for us.
In particular, parts of the documentation were inaccessible to us as beginning 
functional programmers.
We found that documentation with straight-forward examples to be the most 
helpful.

Another component worthy of evaluation is the debugging process.
Debugging in Haskell often takes the form of fixing type errors; which we think 
is a great thing.
However, in the case where debugging extends into runtime behavior, there 
doesn't appear to be any clearly defined approach to effective debugging.
Stack traces are available via profiling, but this was difficult to find and 
as we mentioned, profiling has its own problems.

% - What does the local expert do? Do you have to have one? 
%  
% - To the FP community: Those who want us to abandon our old tools for debugging imperative code have not provided a clear path to follow (both intellectually, and via software tools). 
%  
% - No obvious body of knowledge on code improvement. 
%  
% - The ecosystem contains tools that look promising and work great: Data.Memocombinators, Quickcheck, Parallel Strategies. 
% How to tell what's good? How did we discover what's good? 
% - Profiler can produce backtraces, but this was hard to discover.
